\pagestyle{fancy}
\fancyhead{}

\fancyheadoffset{0cm}
\renewcommand{\headrulewidth}{1pt} 
\renewcommand{\footrulewidth}{0pt}
\fancyhead[L]{{\thesection}}
\fancyhead[L]{\leftmark}

% \fancyhead[LO,LE]{\nouppercase\firstleftmark}
%\fancyhead[RO,RE]{\thesubsection\enspace\subsectiontitle}

\section{Introduction}

\subsection{Epilepsy}

Epilepsy is a chronic noncommunicable neurological disorder characterized as 
the recurrence of unprovoked seizures, which are brief episodes of abnormal
electrical activity in the brain.
Epileptic seizures can result in symptoms such as involuntary movement,
convulsions and loss of consciousness \cite{WHO2016Epilepsy}.

Epilepsy is one of the most common neurological diseases globally: it is
estimated that approximately 50 million people suffer from epilepsy, and
about $5\%$ of people will have experienced at least an unprovoked seizure by
the age of 80 \cite{ep-wilden}.
Approximately $69\%$ of the cases can be controlled using anti-seizure
medications, as drug-resistance is a frequent complication \cite{ep-eadie}.

Seizures consist of the disruption of the normal balance between 
excitatory and inhibitory neurotransmission,
resulting in abnormal, excessive and synchronized neuronal activity.
Epileptic seizures are a highly heterogeneous phenomena.
They can arise from a variety of underlying conditions, and in approximately
$50\%$ of cases, the cause remains unknown \cite{WHO2016Epilepsy}.
Significant variability exists in the type, severity and anatomical origin of a
seizure.
Such differences are present both between different patients and between
different seizures of the same patients \cite{ep-hetero}.  

\subsection{Seizure Onset Zone (SOZ)}

Epileptic seizures are distinguished in two major types: generalized seizures
and focal seizures.
Generalized seizures involve rapid and near-simultaneous abnormal electrical
activity across both cerebral hemispheres, without a stable onset zone. 
Focal seizures initially present electrical symptoms in a specific region of
the brain, called Seizure Onset Zone (SOZ).
Once initiated, the seizure may spread to adjacent areas, involving a vaster 
area of the brain. \cite{ep-class}

At seizure onset, neurons exhibit a transition from normal asynchronous activity
to excessive and synchronized firing.
Seizure activity may manifest as high frequencies oscillations, synchronized
coupling or transient events such as spikes and sharp waves.
These patterns can vary between patients or between different seizures of the 
same patient, but consistently present a transition between normal and 
epileptogenic state.
The SOZ is defined by the brain regions that present the earliest detectable
deviations from the normal background activity of neurons \cite{ep-seiz}.

In drug-resistant focal epilepsy, surgical treatment represents the best chance 
to drastically reduce (or potentially eliminate) the frequency of future
seizures.
Different types of surgery exist, including resective surgery and
radiofrequency thermocoagulation, but all methods aim to remove, destroy, or
functionally disconnect the SOZ from the rest of the brain.
In this context, the accurate localization of the SOZ has major clinical
relevance for pre-surgery planning \cite{ep-sozsurg}.

\subsection{Stereo-electroencephalography}

Because an epileptic seizure is a dynamic transition, SOZ localization requires
a time-resolved measurement of electrical brain activity.
Stereoelectroencephalography (sEEG) is the practice of recording
electroencephalographic signals using surgically implanted electrodes.
Each electrode consists of multiple recording sites, which we will
refer as "contacts", capturing electrical activity from different depths within
the brain.

The strength of sEEG, when compared to electrocorticography (ECoG)
or non-invasive magnetoencephalography (MEG),
is to be able to provide in-depth recordings of electric signals in otherwise
unreachable brain areas. 
However, sEEG is an invasive procedure that requires surgery during the
electrode implatation, therfore it may only be used on patients that do not
respond to medical treatment and who are potential candidates to receive brain
surgery in order to control their symptoms.

\subsection{Challenges and limitations}

Research on SOZ localization is rapidly evolving in recent years.
The increasing availability of high-resolution electrophysiological recordings
and improved computational tools has enabled more sophisticated analyses of
seizure dynamics, progressively transitioning from visual assessment to
quantitative, systematic and reproducible algorithms.
In recent years, machine learning approaches have been increasingly integrated
with graph network theory, as many studies model epileptic seizures as a
network phenomenon involving the mutual relationships in the whole brain
rather than a purely focal event. 
Despite recent methodological advances, major limitations still challenge the
robustness and generalizability of SOZ localization methods \cite{ep-rev}.

Epilepsy is itself a very heterogeneous disorder, presenting both inter-patient
and intra-patient differences in seizure dynamics.
Epilepsy can be caused by a variety of underlying conditions and seizure
patterns depend on the specific cause, which is however only identified in
approximately half of cases \cite{WHO2016Epilepsy}.
Seizures can originate from many brain regions, and exhibit substantially
different behavior depending on the anatomical location of the SOZ.
The performance of machine learning models was shown to heavily depend on the
SOZ location, performing significantly worse on specific regions of the brain
\cite{ep-rev}.
while performing worse on more complex forms of epilepsy.    
The onset area has also been observed to vary across different
seizures within the same patient,
therfore the analysis of a single seizure may lead to an underestimation of the
actual SOZ \cite{ep-hetero2}.

Additionally, the lack of a standardized acquisition protocol makes
comprehensive studies particularly challenging \cite{ep-rev}.
Even within the same acquisition center, the positioning of ECoG or sEEG
sensors is different for each patient, as it is usually conditioned by
preliminar analysis.
The unification of multiple patients in a unique dataset inevitably results in
the inclusion of many brain areas, resulting in highly variable brain signals
in both seizures and resting-state activity.

As in many medical machine learning application, data availability is often a
limitation.
Only a small part of the studies reported in literature uses publicly available
dataset, while the majority of reaserch uses private data with a typically
low number of patients (the average sample size between studies is 23 patients
\cite{ep-rev}).
It has been observed that studies with lower sample size generally report higher
accuracy and sensitivity, while studies with larger sample sizes tend to have 
lower performance. 
This trend suggests that high-performance results may be easier to achieve in
studies with limited data but may not generalize well because of data
heterogeneity \cite{ep-rev2}. 

To evaluate classification performance and assist supervised methods, the list
of resected or thermocoagulated regions is often used as ground truth for the
SOZ localization.
A further complication consists in the reliability of ground truths, as 
long-term postoperative follow-ups are often absent or report insufficient data.
If similar seizures still persist after surgery, it may indicate that the SOZ
reported in the ground truth was inaccurately localized. 
Studies reporting longer follow-up durations may provide more reliable insights 
into the long-term clinical utility of the method \cite{ep-rev}.

Another limitation consists in the high requirement of computational resources.
While deep learning models achieve better performances with respect to 
traditional machine learning methods, high complexity translates in long
computation time.
Network analysis enables to capture complex relationships between different
brain regions, but often requires heavy computation.
The need for computational infrastructure may limit adoption in clinical 
settings, particularly in resource-constrained environments \cite{ep-rev}. 

\subsection{Thesis objective}

The objective of this thesis is to develop an artificial intelligence pipeline
for SOZ localization of multiple patients affected by focal epilepsy.
We aim to design a patient-independent algorithm which is able to distinguish 
between resting-state and epileptogenic activity in sEEG recordings in order to
classify each brain region as SOZ or healthy tissue.

The proposed method combines network theory, deep learning, anomaly detection
and classification approaches to localize the SOZ.
Meaningful features are extracted from sEEG time-series capturing signal,
spectral and network patterns to characterize each channel.
This approach combines both single-channel dynamics and the mutual
relationships between different channels.
Resting-state activity of multiple patients is used to train the model to learn
an encoded representation of the expected behavior of a healthy brain.
Epileptogenic activity is detected as a deviation from resting-state patterns,
and the SOZ detection is implemented by the dynamic evolution of the anomaly
level associated with each brain region.

Instead of directly modeling seizure activity, which is highly heterogeneous
both across and within patients, the proposed framework focuses on learning
the statistical structure of resting-state brain activity, which is more
abundant and less variable than seizure dynamics.
A Variational Autoencoder (VAE) performs non-linear dimensionality reduction
to construct a smooth, interpolable, lower-dimensional encoded representation
of the joint distribution of features.
SOZ localization can be formulated as the identification of statistical
anomalies in the encoded distribution of resting-state features: pathological
dynamics are identified as outliers in the expected resting-state distribution.
This approach is purposly unsupervised, to provide robustness against possible
ground truth inaccuracies.

A key objective in this work is the development of a patient-independent model.
We aim to select a unique set of parameters to produce an accurate SOZ
classification on multiple patients, as many existing methods are tailored to
single patients and require manual tuning or individualized calibration
\cite{ep-rev2}.
While effective in controlled settings, these strategies limit scalability and
automation.

The proposed method uses an automatic pipeline to capture common 
statistical properties of sEEG signals despite the individual differences
between multiple patients.
Hyperparameter optimization is performed to select a unique set of parameters
for all patients, and cross-validation was used to estimate generalization
capabilites on unseen patients.
The model is then evaluated on an independent set of test patients, reporting
established classification metrics.

Particular attention was paid to avoid information leakage: patients used for 
designing the algorithm were never included in validation or test sets.
Similarly, test patients were completely unseen during both training and
validation, and no data from these patients was used at any stage of model
development.
Such separation ensures that the reported performance reflects, without
overestimating, the true generalization capabilites of the model.