\pagebreak
    \hspace{0pt}
    \vfill
    \addcontentsline{toc}{section}{Abstract}
    \section*{Abstract}
    
    In drug-resistant focal epilepsy, surgical treatment represents the best
    chance to reduce seizure frequency, and accurate localization of the
    Seizure Onset Zone (SOZ) is critical for pre-surgery planning.
    However, the high inter-patient heterogeneity of stereo-electroencephalography
    (sEEG) recordings makes the development of generalizable, patient-independent
    methods particularly challenging.

    This study proposes an automated pipeline for the localization of the SOZ
    across multiple patients using a unified parameter set.
    23 patients were included in the dataset, including resting-state
    and seizure sEEG recordings.
    22 features are extracted from data:
    signal features characterizing the time domain, spectral features capturing
    information from the frequency domain and network features modelling the
    synchronous relationships between brain regions.

    A Variational Autoencoder (VAE) produces a latent representation of the
    resting-state features, subsequently a Gaussian Mixture Model (GMM)
    fits the encoded distribution and scores the likelihood of seizure features.
    A classifier decides which brain regions compose the SOZ based on the
    negative log-likelihood of seizure features, which is used as an indicator
    for anomaly detection.
    Hyperparameter optimization is performed in order to tune the 10
    hyperparameters that define the algorithm.
    Optimization includes a preliminary validation phase using Leave-One-Out
    Cross-Validation (LOOCV), then the final model is trained on the full
    training set and evaluated on an independent test set of unseen patients.

    The proposed model achieved promising results during cross-validation,
    however a significant performance drop was observed on unseen test patients,
    reflecting the intrinsic heterogeneity of sEEG recordings across subjects.
    A comparison with an automated version of the Epileptogenicity Index (EI),
    a well-established and widely used quantitative measure for SOZ localization,
    demonstrates that the proposed method achieves improved performance,
    highlighting limitations in the traditional EI approach.

    Results highlight that inter-patient variability remains the primary
    challenge in SOZ localization, however potential strategies are proposed
    to direct future work towards patient-adaptive approaches.
    \vfill
    \hspace{0pt}
\pagebreak

